---
title: "UNHCR Funding Analysis  "
subtitle: "`r paste('Donor Profile:', params$name)`"
author: "AI Generated Analysis based on the open data shared publicly by UNHCR as part of the [International Aid Transparency Initiative (IATI)](https://www.unhcr.org/iati-international-aid-transparency-initiative). Beware of data limitations and potential hallucinations! Thanks for reporting any [issues here](https://github.com/Edouard-Legoupil/iati/issues/new) -- [View all Reports](../articles/ai-powered-reports.html)"
format:
  html:
    toc: true
    code-fold: true
    self-contained: true

params:
  name: "Private donors"

  year: null
  start_year: null
  end_year: null
  ctr_name: null

  top_n_countries: 10
  top_n_interest: 10
  other_stat: "mean"
  funding_by: "global"
  earmarking_by: "region"
  volatility_time_unit: "quarter"
  volatility_forecast_h: 8
  volatility_show_ci: true

  # strict one page tuning
  fig_width: 5.55
  fig_height: 1.75
  dpi: 260

  # typography
  base_text_size: 26
  axis_text_size: 18
  legend_text_size: 18
  strip_text_size: 21
  plot_family: "Lato"           # will auto-fallback to sans if not available to R device

  # AI narration
  use_ai: true
  ai_provider: "azure"          # azure|auto
  ai_model: "gpt-4.1-mini"
  max_tokens_short_desc: 30        # only used as instruction in prompt
  max_tokens_long_desc: 500
  ai_max_tokens_exec: 100         # only used as instruction in prompt
  ai_debug: true
---

```{r setup, include=FALSE}
# Set default Quarto options
knitr::opts_chunk$set(echo = FALSE, fig.cap = TRUE, warning = FALSE, message = FALSE)
# Load necessary libraries
suppressPackageStartupMessages({
  library(dplyr)
  library(ggplot2)
  library(scales)
  library(lubridate)
  library(stringr)
  library(unhcrthemes)
  library(iati)
  library(ellmer)
})

 

normalize_year <- function(y) if (is.null(y) || length(y) == 0) NULL else as.integer(y)
normalize_chr  <- function(x) if (is.null(x) || length(x) == 0) NULL else as.character(x)

# -------------------------
# AI narration (ellmer) 
# -------------------------
ai_enabled <- isTRUE(params$use_ai)
ai_debug   <- isTRUE(params$ai_debug)

ai_log <- character()
ai_note <- function(x) { ai_log <<- c(ai_log, x) }

# ---- 

p_year  <- normalize_year(params$year)
p_start <- if (is.null(params$start_year)) NULL else as.integer(params$start_year)
p_end   <- if (is.null(params$end_year)) NULL else as.integer(params$end_year)
p_ctr   <- normalize_chr(params$ctr_name)

# Determine a safe font family for ggplot rendering
plot_family_safe <- params$plot_family
if (!is.null(plot_family_safe) && plot_family_safe == "Lato") {
  ok <- FALSE
  if (requireNamespace("systemfonts", quietly = TRUE)) {
    mf <- tryCatch(systemfonts::match_font("Lato"), error = function(e) NULL)
    ok <- !is.null(mf) && !is.null(mf$path) && nzchar(mf$path)
  }
  if (!ok) plot_family_safe <- "sans"
}

placeholder_plot <- function(title = "No data", subtitle = "") {
  ggplot2::ggplot() +
    ggplot2::theme_void() +
    ggplot2::annotate("text", x = 0, y = 0.1, label = title, size = 5, fontface = "bold") +
    ggplot2::annotate("text", x = 0, y = -0.1, label = subtitle, size = 3.5) +
    ggplot2::xlim(-1,1) + ggplot2::ylim(-1,1)
}

safe_plot <- function(expr) {
  res <- tryCatch(expr, error = function(e) placeholder_plot("Plot unavailable", conditionMessage(e)))
  if (is.null(res)) placeholder_plot("No data for this view", "Function returned NULL") else res
}

render_plot <- function(x) {
  if (inherits(x, "ggplot")) print(x)
  else if (inherits(x, c("grob", "gtable"))) { grid::grid.newpage(); grid::grid.draw(x) }
  else print(placeholder_plot("Unsupported plot object", class(x)[1]))
}

normalize_plot <- function(p) {
  if (!inherits(p, "ggplot")) return(p)
  p + ggplot2::theme(
    text = ggplot2::element_text(size = params$base_text_size, family = plot_family_safe),
    axis.text = ggplot2::element_text(size = params$axis_text_size, family = plot_family_safe),
    axis.title = ggplot2::element_text(size = params$axis_text_size, family = plot_family_safe),
    legend.text = ggplot2::element_text(size = params$legend_text_size, family = plot_family_safe),
    legend.title = ggplot2::element_text(size = params$legend_text_size, family = plot_family_safe),
    strip.text = ggplot2::element_text(size = params$strip_text_size, family = plot_family_safe),
    plot.margin = ggplot2::margin(1, 1, 1, 1, unit = "pt"),
    legend.position = "bottom",
    legend.key.height = ggplot2::unit(6, "pt"),
    legend.key.width  = ggplot2::unit(10, "pt")
  )
}

chat =  ellmer::chat_azure_openai(
  system_prompt = "You are a humanitarian funding analyst.Your task is to build a concise executive summary to describe the profile of a specific donor based on provided information",
  model = params$ai_model,
  api_version = Sys.getenv("AZURE_OPENAI_API_VERSION"),
  endpoint = Sys.getenv("AZURE_OPENAI_ENDPOINT"),
  api_key = Sys.getenv("AZURE_OPENAI_API_KEY"))




 
```

```{r summary}
#| message: false
#| warning: false
#| include: false

# Plots
donor_ranking_raw <- safe_plot(
  show_donor_ranking(
    donor_name = params$name,
    year = if (is.null(p_year))
      NULL
    else
      max(p_year),
    top_n_display = 10,
    ctr_name = p_ctr,
    verbose = FALSE
  )
)

donor_regional_focus_treemap_raw <- safe_plot(
  plot_donor_regional_focus_treemap(
    donor_name = params$name,
    year = 2025,
    label_font_size = 5
  )
)



donor_funding_over_time_raw <- safe_plot(
  show_donor_funding_over_time(donor_name = params$name, 
                               year = c(2022, 2023, 20234, 2025),
                               by = "region"))

donor_activity_diversification_raw <- safe_plot(
  show_donor_activity_diversification(
    donor_name = params$name, 
    start_year = p_start, 
    end_year = p_end,
    other_stat = params$other_stat, 
    verbose = FALSE))

donor_earmarking_raw <- safe_plot(
  show_donor_earmarking(
    donor_name = params$name, 
    year = c(  2025),
    by = params$earmarking_by,
    ctr_name = p_ctr))

# donor_transaction_desc_raw <- safe_plot(
#   show_donor_transaction_desc(donor_name = params$name, 
#                               by_primary = "transaction_description", 
#                               by_secondary = "unhcr_region",
#                               year = p_year, 
#                               ctr_name = p_ctr, 
#                               top_n = params$top_n_interest))

donor_funding_volatility_raw <- safe_plot(
  show_donor_funding_volatility(donor_name = params$name,
  year = 2023:2026,
  forecast_horizon = 8,
  time_unit = "month",
  verbose = FALSE))

compare_donor_profiles_raw <- safe_plot(            
    compare_donor_profiles(  donor_names =  params$name,
                              by = "earmarking",
                              year = p_year,
                              avg_method = "pooled",
                              display_mode = "relative"))

compare_donor_profiles2_raw <- safe_plot(            
    compare_donor_profiles(  donor_names =  params$name,
                              by = "unhcr_region",
                              year = p_year,
                              avg_method = "pooled",
                              display_mode = "relative"))


donor_geographic_priority_shift_raw <- safe_plot(    
  show_donor_geographic_priority_shift(  donor_name =  params$name,
                                         top_n_countries = 6,
                                         start_year = 2022,
                                         end_year = 2025))

              



stories <- c(
  "Ranking" = generate_plot_story(donor_ranking_raw, 
                                  provider = params$ai_provider, model = params$ai_model,
                                  max_tokens_short_desc = params$max_tokens_short_desc,
                                  max_tokens_long_desc= params$max_tokens_long_desc ),
  
  "Focus" = generate_plot_story( donor_regional_focus_treemap_raw, 
                                  provider = params$ai_provider, model = params$ai_model,
                                  max_tokens_short_desc = params$max_tokens_short_desc,
                                  max_tokens_long_desc= params$max_tokens_long_desc ),
  
  "Funding" = generate_plot_story(donor_funding_over_time_raw, 
                                  provider = params$ai_provider, model = params$ai_model,
                                  max_tokens_short_desc = params$max_tokens_short_desc,
                                  max_tokens_long_desc= params$max_tokens_long_desc),
  
  "Diversification" = generate_plot_story(donor_activity_diversification_raw, 
                                  provider = params$ai_provider, model = params$ai_model,
                                  max_tokens_short_desc = params$max_tokens_short_desc,
                                  max_tokens_long_desc= params$max_tokens_long_desc),
  
  "Earmarking" = generate_plot_story(donor_earmarking_raw, 
                                  provider = params$ai_provider, model = params$ai_model, 
                                  max_tokens_short_desc = params$max_tokens_short_desc,
                                  max_tokens_long_desc= params$max_tokens_long_desc),
  
  # "Interest" = generate_plot_story(donor_transaction_desc_raw, 
  #                                 provider = params$ai_provider, model = params$ai_model, 
  #                                 max_tokens = params$ai_max_tokens_exec),
  
  
  "Volatility" = generate_plot_story(donor_funding_volatility_raw, 
                                  provider = params$ai_provider, model = params$ai_model, 
                                  max_tokens_short_desc = params$max_tokens_short_desc,
                                  max_tokens_long_desc= params$max_tokens_long_desc),
  
  
  "Profile" = generate_plot_story(compare_donor_profiles_raw, 
                                  provider = params$ai_provider, model = params$ai_model, 
                                  max_tokens_short_desc = params$max_tokens_short_desc,
                                  max_tokens_long_desc= params$max_tokens_long_desc),
  
  
  "ProfileReg" = generate_plot_story(compare_donor_profiles2_raw, 
                                  provider = params$ai_provider, model = params$ai_model, 
                                  max_tokens_short_desc = params$max_tokens_short_desc,
                                  max_tokens_long_desc= params$max_tokens_long_desc),
  
  "Shift" = generate_plot_story(donor_geographic_priority_shift_raw, 
                                  provider = params$ai_provider, model = params$ai_model, 
                                  max_tokens_short_desc = params$max_tokens_short_desc,
                                  max_tokens_long_desc= params$max_tokens_long_desc) 
)


chat =  ellmer::chat_azure_openai(
  system_prompt = "You are a humanitarian funding analyst.Your task is to build a concise executive summary to describe the profile of a specific recipient based on provided information, Your goal is to highlight the most relevant information to inform fundraising activities and project design. ",
  model = "gpt-4.1-mini",
  api_version = Sys.getenv("AZURE_OPENAI_API_VERSION"),
  endpoint = Sys.getenv("AZURE_OPENAI_ENDPOINT"),
  api_key = Sys.getenv("AZURE_OPENAI_API_KEY"))

exec <- chat$chat( paste0(" Build a short ",params$ai_max_tokens_exec , "  words Donor Profile for ",
                           params$name, ". \n. Ground your summary on the below information:",
                           stories["Ranking.long_desc"] , "\n\n",
                           stories["Focus.long_desc"] , "\n\n",
                           stories["Funding.long_desc"] , "\n\n",  
                           stories["Diversification.long_desc"] , "\n\n", 
                           stories["Earmarking.long_desc"] , "\n\n",
                           stories["Profile.long_desc"] , "\n\n",
                           stories["ProfileReg.long_desc"] , "\n\n",
                           stories["Shift.long_desc"] , "\n\n",    
                           stories["Volatility.long_desc"] , "\n\n", 
                          " \n\n INSTRUCTIONS: Do not repeat the provided content - focus on extracting key high level information for a fund raisers audience"
                           )) 
```

## Executive Summary

`r exec`

## Ranking

`r stories["Ranking.long_desc"]`
  
```{r Ranking}
#| fig.retina: 2.0
#| fig.width: 8.0
#| fig.asp: 0.618
#| fig.align: center
#| out.width: 90%

donor_ranking <- normalize_plot(donor_ranking_raw + 
                    ggplot2::labs(subtitle = stories["Ranking.short_desc"]))
donor_ranking
```


## Focus Portfolio

`r stories["Focus.long_desc"]`  
  
```{r Focus}
#| fig.retina: 2.0
#| fig.width: 8.0
#| fig.asp: 0.618
#| fig.align: center
#| out.width: 90%

donor_regional_focus_treemap <- normalize_plot(donor_regional_focus_treemap_raw +
                ggplot2::labs(subtitle = stories["Focus.short_desc"]))

donor_regional_focus_treemap

plot_donor_location_map(
  donor_name = params$name,
  year = 2025,
  top_n_locations = 5,
  max_symbol_size = 20
)
```

## Funding  Trends

`r stories["Funding.long_desc"]`  
  
```{r Funding}
#| fig.retina: 2.0
#| fig.width: 8.0
#| fig.asp: 0.618
#| fig.align: center
#| out.width: 90%

donor_funding_over_time <- normalize_plot(donor_funding_over_time_raw +
                ggplot2::labs(subtitle = stories["Funding.short_desc"]))

donor_funding_over_time
```



## Earmarking Behavior

`r stories["Earmarking.long_desc"]`  
  
```{r Earmarking}
#| fig.retina: 2.0
#| fig.width: 8.0
#| fig.asp: 0.618
#| fig.align: center
#| out.width: 90%

donor_earmarking <- normalize_plot(donor_earmarking_raw + 
              ggplot2::labs(subtitle = stories["Earmarking.short_desc"]))
donor_earmarking
```

## Comparison with other donors

`r stories["Profile.long_desc"]`  
  
```{r Profile}
#| fig.retina: 2.0
#| fig.width: 8.0
#| fig.asp: 0.618
#| fig.align: center
#| out.width: 90%

compare_donor_profiles <- normalize_plot(compare_donor_profiles_raw +
                   ggplot2::labs(subtitle = stories["Profile.short_desc"]))

compare_donor_profiles
```

`r stories["ProfileReg.long_desc"]`  
  
```{r Profile2}
#| fig.retina: 2.0
#| fig.width: 8.0
#| fig.asp: 0.618
#| fig.align: center
#| out.width: 90%

compare_donor_profiles2 <- normalize_plot(compare_donor_profiles2_raw +
                   ggplot2::labs(subtitle = stories["ProfileReg.short_desc"]))

compare_donor_profiles2
```


## Main Activitiies Shift

`r stories["Shift.long_desc"]`  
  
```{r Shift}
#| fig.retina: 2.0
#| fig.width: 8.0
#| fig.asp: 0.618
#| fig.align: center
#| out.width: 90%


donor_geographic_priority_shift <- normalize_plot(
  donor_geographic_priority_shift_raw + 
    ggplot2::labs(subtitle = stories["Shift.short_desc"]))

donor_geographic_priority_shift
```

## Activities Diversification 

`r stories["Diversification.long_desc"]`  
  
```{r Diversification}
#| fig.retina: 2.0
#| fig.width: 8.0
#| fig.asp: 0.618
#| fig.align: center
#| out.width: 90%

donor_activity_diversification <- normalize_plot(
  donor_activity_diversification_raw +
    ggplot2::labs(subtitle = stories["Diversification.short_desc"]))

donor_activity_diversification
```



## Transaction Volatility

`r stories["Volatility.long_desc"]`  
  
```{r Volatility}
#| fig.retina: 2.0
#| fig.width: 8.0
#| fig.asp: 0.618
#| fig.align: center
#| out.width: 90%


donor_funding_volatility <- normalize_plot(donor_funding_volatility_raw +
                ggplot2::labs(subtitle = stories["Volatility.short_desc"]))
donor_funding_volatility
```


